Supplementary Methods for\\ \emph{Effectiveness of UNAIDS targets and HIV vaccination across 127 countries}

### 2016-11-20 Documents/supplementary_information.tex
Switch from biblatex/biber to bibtex so I can offset the reference numbers.

\usepackage[style=nature]{biblatex}
\addbibresource{supplementary_information.bib}

@@

The model simulation and analysis tools, written in Python, are
publicly available \autocite{medlock2016-git}.

@@

We developed a country-level continuous-time mathematical model that
divides people aged $15$\;y and older into non-overlapping HIV states.
We focused on this age group because it accounts for the vast majority
of global PLHIV \autocite{UNICEF}, 93\% in 2014; because
population-level HIV estimates and demographic data are typically only
estimated for ages $15$--$49$; and to capture the continuing burden of
infections in people ages $50$ and older.  Our model has 8 HIV states
(Extended Data Fig.~\ref*{model_diag}): susceptible to HIV infection
($S$ is the number of people susceptible), vaccinated against HIV
($R$), acute HIV infection ($A$), undiagnosed HIV infection ($U$),
diagnosed but untreated HIV infection ($D$), treated without viral
suppression ($T$), viral suppression ($V$), and having AIDS ($W$).
Viral suppression is defined as a viral load of less than 1000 viral
RNA copies per mL of blood.  Viral suppression not only extends
survival but also dramatically reduces transmission
\autocite{vernazza2000, ioannidis2001, attia2009, May2014-gp}.
Simulations of our model project the number of people in each HIV
stratum from 2015 to 2035.

@@

The background mortality rate ($\mu$) in each of the 127 countries was
collected from published estimates
\autocite{World_Development_Indicators2013-ee}.  The recruitment rate
($\kappa$) into the age $15\;\text{y}$ and older group was set by the
difference between the country's population growth rate
\autocite{WorldBankpg} and its background mortality rate
\autocite{World_Development_Indicators2013-ee}, as would be true if
the population was at its stable age structure.  (More on these data
sources is in \autoref{data_sources}.)

\begin{table}
  \begin{center}
    \begin{tabularx}{\textwidth}{lXlll}
      \hline
      & Definition & Value & Reference \\
      \hline
      $\mu$ & Background death rate
      & Country specific, Supplementary Table 2
      & \autocite{World_Development_Indicators2013-ee} \\
      $\kappa$ & Recruitment rate
      & Country specific, Supplementary Table 2
      & \autocite{World_Development_Indicators2013-ee, WorldBankpg} \\
      $\delta$	& Rate of leaving acute infection
      & $\Triangular$(2, 4.14, 9.6)\;y$^{-1}$
      & \autocite{Hollingsworth2008-iy} \\
      $\sigma$	& Rate of developing AIDS without viral suppression
      & 0.1064\;y$^{-1}$ & \autocite{Morgan2002-cq} \\
      $\theta$ & Rate of developing AIDS with viral suppression
      & Eq.~(\ref{theta}) & --- \\
      $\gamma$ & Rate of viral suppression
      & $\Uniform$(0.5, 1.5)\;y$^{-1}$
      & \autocite{Currie2009-yz} \\
      $\nu$ & Death rate from AIDS & 0.5\;y$^{-1}$
      & \autocite{Morgan2002-cq} \\
      $\rho$ & Vaccine efficacy & 50\%\,[30\%, 70\%] & --- \\
      $\omega$	& Reduction in lifetime with viral suppression
      & $\Uniform$(5, 8)\;y
      & \autocite{Samji2013-kf, Unaids2014-ue} \\
      $\tau_{A}$ & Transmissibility during acute phase
      & $\Triangular$(0.0039, 0.0082, 0.0150)
      & \autocite{Wawer2005-us, Skarbinski2015-ni}\\
      $\tau_{U}$ & Transmissibility after acute phase
      & $\Triangular$(0.00077, 0.0014, 0.00251)
      & \autocite{Hughes2012-so} \\
      $\varepsilon$ & Relative transmissibility with
      viral suppression & $\BetaPERT$(0.08, 0.002, 0.57)
      & \autocite{Donnell2010-xo, attia_2009, wilson_2012, jia_2013,
        rodger_2016} \\
      $n$ & Coital acts per year & $\Uniform$(96, 108)
      & \autocite{Wawer2005-us, Abdool_Karim2010-cm}\\
      $\alpha$ & Diagnosis rate & See eq.~(\ref{diagnosis_rate}) & --- \\
      $\phi$ & Treatment rate & See eq.~(\ref{treatment_rate}) & --- \\
      $\psi$ & Rate of relapse to untreated & See eq.~(\ref{relapse_rate})
      & --- \\
      $\zeta$ & Vaccination rate & See eq.~(\ref{vaccination_rate}) & --- \\
      \hline
      \vspace*{-2.5em}
    \end{tabularx}
    \caption{Model parameters. $\Triangular$, $\Uniform$, and
      $\BetaPERT$ are sampling distributions (\autoref{uncertainty}).
      For vaccine efficacy, $50\%$ was the baseline, and $30\%$ and
      $70\%$ were also used in a scenario analysis.}
    \label{model_param}
  \end{center}
\end{table}

People susceptible to HIV ($S$) progress to acute infection ($A$)
according to the force of infection ($\lambda$), which depends on the
HIV transmission rate, estimated from the country-specific incidence
and prevalence data (\autoref{model_fitting}), adjusted for the
current number of acutely infected, untreated or ineffectively
treated, and viral suppression PLHIV.  Acute infection is
characterized by high viral loads and lasts an average of 2.90 [95\%
confidence interval 1.23, 6.00]\;months
\autocite{Hollingsworth2008-iy}.  (We converted these durations to
rates like $1 / 2.90\;\text{month$^{-1}$} \approx
4.14\;\text{y$^{-1}$}$ and set the parameter using random samples from
$\Triangular(2, 4.14, 9.6)\;\text{y$^{-1}$}$ to capture the
uncertainty in its value.  See \autoref{uncertainty} for the
definition of $\Triangular(a, b, c)$ and other distributions.)  We
assumed that people remain undiagnosed during their acute-infection
phase. After acute infection, people move into the chronic undiagnosed
class ($U$).  People in the undiagnosed class will be diagnosed at the
rate $\alpha$, which is determined by the country's current diagnosis
level and its diagnosis target (\autoref{targets}).  Diagnosed people
($D$) transition to the treated compartment ($T$) upon starting ART at
the rate $\phi$, dependent on the country's current treatment level
and its treatment target (\autoref{targets}).  After some time,
treated people then transition to the viral-suppression class ($V$),
at the rate $\gamma$.  We took $\gamma$ as a random sample from
$\Uniform(0.5, 1.5)\;\text{y$^{-1}$}$, that is, viral suppression
occurs after between 8 months and 2 years on treatment
\autocite{Currie2009-yz}.  Disengagement from treatment moves people
from treated ($T$) and viral suppression ($V$) at the rate $\psi$,
dependent on the current level of viral suppression and the country's
target (\autoref{targets}), back to the diagnosed class ($D$).

PLHIV without viral suppression (compartments $U$, $D$, and $T$)
develop AIDS (compartment $W$) after an average duration of 10.4 years
\autocite{Morgan2002-cq}, i.e. with transition rate
$\sigma = 0.1064\;\text{y$^{-1}$}$.  People who have achieved viral
suppression ($V$) develop AIDS at rate $\theta$, slower than PLHIV
without viral suppression.  People with viral suppression are
estimated to live $5$--$8$\;y shorter lives than the non-HIV infected
\autocite{Samji2013-kf, Unaids2014-ue}, which we quantified by
randomly sampling the parameter $\omega$ from
$\Uniform(5, 8)\;\text{y}$.  Assuming no relapse to untreated, the
duration of viral suppression is
\begin{equation}
  \frac{1}{\theta + \mu} = \frac{1}{\mu} - \omega - \frac{1}{\nu}.
\end{equation}
The term to the left of the equals sign is the time until leaving
viral suppression by either progression to AIDS (at rate $\theta$) or
background mortality (at rate $\mu$).  The term to the right of the
equals sign is the lifespan of the non-HIV infected, minus the
reduction in life due to having HIV with viral suppression, minus the
duration of the AIDS stage.  Solving for $\theta$ gives the rate of
progression to AIDS with viral suppression as
\begin{equation}
  \label{theta}
  \theta = \frac{1}{1/\mu - \omega - 1/\nu} - \mu.
\end{equation}

@@

The relative transmission rates of acute, unsuppressed, and suppressed
infected people are
\begin{equation}
  \label{betas}
  \begin{split}
    \beta_A &= 1 - (1 - \tau_A)^n,
    \\
    \beta_U &= 1 - (1 - \tau_U)^n,
    \\
    \beta_V &= 1 - (1 - \varepsilon \tau_U)^n,
  \end{split}
\end{equation}
where the $\tau$'s are the relative transmissibilities of people with
acute ($A$), unsuppressed ($U$ and $T$), and suppressed ($V$)
infections, $n$ is the annual number of coital acts, and $\varepsilon$
is the relative transmissibility with viral suppression.  For
transmissibility during the acute phase, we sampled $\tau_A$ from
$\Triangular$(0.0039, 0.0082, 0.0150)
\autocite{Wawer2005-us, Skarbinski2015-ni}, whereas, for transmissibility
for unsuppressed people, we sampled $\tau_U$ from
$\Triangular(0.00077, 0.0014, 0.00251)$ \autocite{Hughes2012-so}.  The
transmissibility of people with viral suppression relative to those
without suppression has been estimated at 0.08 (95\% confidence
interval 0.002, 0.57) \autocite{Donnell2010-xo}: due to the extremely long
left tail, we chose to sample the relative transmissibility
$\varepsilon$ from $\BetaPERT(0.08, 0.002, 0.57)$ rather than a
$\Triangular$ distribution.  (See \autoref{uncertainty} for a full
description of $\BetaPERT(a, b, c)$.)  The annual number of coital
acts $n$ was sampled from $\Uniform(96, 108)$
\autocite{Wawer2005-us, Abdool_Karim2010-cm}.  The country-specific
transmission rate $\eta$ was derived from UNAIDS estimates of each
country's longitudinal HIV prevalence and incidence
(\autoref{model_fitting}).

@@

The parametrized differential equations were solved numerically from
year 2015 to year 2035 using the LSODA routine
\autocite{odepack, scipy, medlock2016-git} and four effectiveness
outcomes were computed from the solutions.
\begin{description}
\item[New infections] New infections from 2015 to time $t$ are given
  by $Y(t)$.

@@

For 127 countries, we found sufficient published estimates and data to
parametrize our model.  Demographic data, including population growth
rate \autocite{WorldBankpg}, death rate
\autocite{World_Development_Indicators2013-ee}, and number of people
aged 15--49 years \autocite{The_World_Bank2016-fd} were obtained from
the World Bank (Supplementary Table 2). Longitudinal HIV prevalence
(Supplementary Table 3) and incidence (Supplementary Table 4)
estimates for ages 15--49, spanning from as early as 1990 for some
countries to 2015, were primarily obtained from the AIDSinfo database
produced by UNAIDS \autocite{Unaids2016-an}. Other sources included
UNAIDS Country Progress Reports \autocite{Unaids2016-am} published
from 2012 to 2016 and AIDS Data Hub \autocite{AIDSdatahub-fg}, which
compiles data from UNAIDS, UNICEF, WHO, and the Asian Development
Bank. These sources were also used to inform the initial conditions
for the model: the number of people diagnosed with HIV, the number on
ART, and the number who have viral suppression or have been retained
on treatment for at least 12 months (Supplementary Table 5). Where
estimates or data were not available from these sources, alternative
resources such as peer-reviewed journal articles and country health
ministry reports were consulted.  See Supplementary Tables 2--5 for
full information on the sources used for each country.  These source
tables are also available in the public source-code repository
\autocite{medlock2016-git}.

@@

Using available historical country-level estimates of prevalence and
incidence (Supplementary Tables 3 and 4), and published estimates of
transmissibility
\autocite{Wawer2005-us,Donnell2010-xo,Hughes2012-so,Skarbinski2015-ni},
we derived an average country-specific transmission rate, which then
informs the acute, unsuppressed, and suppressed transmission rates.
This calibration method provided a recent estimate of country-specific
transmission, while using historical data to smooth out short-term
variations.

@@

To reflect the uncertainty in $\beta(t)$, we assumed that it follows a
lognormal distribution (\autoref{uncertainty}). For the parameters
$\mu$ and $\sigma^2$ of the lognormal distribution we used the
exponentially weighted mean and variance \autocite{holt2004} over time
(with half-life 1\;y) of $\log \beta(t)$ for the year 2015.  That is,
\begin{equation}
  \label{lognormal_params}
  \begin{split}
    \mu &= \frac{\sum_{i = 1}^n  2^{t_i}
      \log \beta(t_i)}
    {\sum_{i = 1}^n 2^{t_i}},
    \\
    \sigma^2 &= \frac{\big(\sum_{i = 1}^n 2^{t_i}\big)
      \left[\sum_{i = 1}^n  2^{t_i}
        \big(\log \beta(t_i) - \mu\big)^2\right]}
    {\big(\sum_{i = 1}^n 2^{t_i}\big)^2
      - \sum_{i = 1}^n 2^{2 t_i}}.
  \end{split}
\end{equation}
A half-life of 1\;y means that the estimate of the transmission rate
from year $i$ is weighted half as much as the estimate from year
$i + 1$, a quarter as much as the estimate from year $i + 2$, and so
on.  This approach captured the recent country-level transmission
behavior while still allowing for the use of less-recent estimates to
smooth short-term variations (\autoref{transmission_rate}).

@@

Projections were generated for each country through model simulations
using 1000 random samples from published parameter distributions
(\autoref{model_param}) and the country-specific estimated
distribution of the transmission rate (\autoref{model_fitting}). The
parameters were sampled using Latin hypercube sampling
\autocite{blower1994}.  We also used the variance reduction technique of
running different interventions with the same parameter samples, i.e.
one set of sample parameter values is drawn, status quo and
intervention scenarios are all run with those sample values and
compared, and this process is repeated for each of the 1000 sample
sets of parameters \autocite{shechter2006}.

The model outcomes were summarized using the median, 1st and 3rd
quartiles (i.e. 25th and 75th percentiles), and 5th and 95th
percentiles.  We calculated partial rank correlation coefficients
(PRCC) \autocite{blower1994} to measure the independent effect of each
non-vaccine parameter on model projections of global new HIV
infections from 2015 to 2035 (Extended Data Fig.~\ref*{PRCCs}).

@@

\item[Beta-PERT] $\BetaPERT(a, b, c)$ is the Beta-PERT probability
  distribution \autocite{malcom1959} with minimum $a$, mode $b$, and
  maximum $c$.  It has density function
  \begin{equation}
    \label{BetaPERT}
    f_{\BetaPERT}(x) =
    \begin{cases}
      \frac{(x - a)^{v - 1} (c - x)^{w - 1}}{(c - a)^{v + w - 2} B(v, w)}
      & \text{if $a \leq x \leq c$,}
      \\
      0 & \text{otherwise,}
    \end{cases}
  \end{equation}
  where
  \begin{equation}
    \begin{split}
      \mu &= \frac{a + \lambda b + c}{\lambda + 2},
      \\
      v &= \frac{(\mu - a)(2 b - a - c)}{(b - \mu) (c - a)},
      \\
      w &= \frac{v (a - \mu)}{\mu - c},
    \end{split}
  \end{equation}
  and $B(v, w)$ is the standard beta function \autocite{davis1972}.
  % \cite[\S6.2]{davis1972}
  We used the standard value of $\lambda = 4$.

@@

\printbibliography

>>>>

% \usepackage[style=nature]{biblatex}
% \addbibresource{supplementary_information.bib}
\usepackage[sort&compress]{natbib}
\usepackage{hypernat}
\citestyle{nature}

% For Nature, add 27 to reference numbers to start after
% last number in manuscript.
\usepackage{etoolbox}
\makeatletter
\apptocmd{\thebibliography}{\global\c@NAT@ctr 27\relax}{}{}
\makeatother

@@

The model simulation and analysis tools, written in Python, are
publicly available\cite{medlock2016-git}.

@@

We developed a country-level continuous-time mathematical model that
divides people aged $15$\;y and older into non-overlapping HIV states.
We focused on this age group because it accounts for the vast majority
of global PLHIV\cite{UNICEF}, 93\% in 2014; because population-level
HIV estimates and demographic data are typically only estimated for
ages $15$--$49$; and to capture the continuing burden of infections in
people ages $50$ and older.  Our model has 8 HIV states (Extended Data
Fig.~\ref*{model_diag}): susceptible to HIV infection ($S$ is the
number of people susceptible), vaccinated against HIV ($R$), acute HIV
infection ($A$), undiagnosed HIV infection ($U$), diagnosed but
untreated HIV infection ($D$), treated without viral suppression
($T$), viral suppression ($V$), and having AIDS ($W$).  Viral
suppression is defined as a viral load of less than 1000 viral RNA
copies per mL of blood.  Viral suppression not only extends survival
but also dramatically reduces
transmission\cite{vernazza2000, ioannidis2001, attia2009, May2014-gp}.
Simulations of our model project the number of people in each HIV
stratum from 2015 to 2035.

@@

The background mortality rate ($\mu$) in each of the 127 countries was
collected from published
estimates\cite{World_Development_Indicators2013-ee}.  The recruitment
rate ($\kappa$) into the age $15\;\text{y}$ and older group was set by
the difference between the country's population growth
rate\cite{WorldBankpg} and its background mortality
rate\cite{World_Development_Indicators2013-ee}, as would be true if
the population was at its stable age structure.  (More on these data
sources is in \autoref{data_sources}.)

\begin{table}
  \begin{center}
    \begin{tabularx}{\textwidth}{lXlll}
      \hline
      & Definition & Value & Reference \\
      \hline
      $\mu$ & Background death rate
      & Country specific, Supplementary Table 2
      & \cite{World_Development_Indicators2013-ee} \\
      $\kappa$ & Recruitment rate
      & Country specific, Supplementary Table 2
      & \cite{World_Development_Indicators2013-ee, WorldBankpg} \\
      $\delta$	& Rate of leaving acute infection
      & $\Triangular$(2, 4.14, 9.6)\;y$^{-1}$
      & \cite{Hollingsworth2008-iy} \\
      $\sigma$	& Rate of developing AIDS without viral suppression
      & 0.1064\;y$^{-1}$ & \cite{Morgan2002-cq} \\
      $\theta$ & Rate of developing AIDS with viral suppression
      & Eq.~(\ref{theta}) & --- \\
      $\gamma$ & Rate of viral suppression
      & $\Uniform$(0.5, 1.5)\;y$^{-1}$
      & \cite{Currie2009-yz} \\
      $\nu$ & Death rate from AIDS & 0.5\;y$^{-1}$
      & \cite{Morgan2002-cq} \\
      $\rho$ & Vaccine efficacy & 50\%\,[30\%, 70\%] & --- \\
      $\omega$	& Reduction in lifetime with viral suppression
      & $\Uniform$(5, 8)\;y
      & \cite{Samji2013-kf, Unaids2014-ue} \\
      $\tau_{A}$ & Transmissibility during acute phase
      & $\Triangular$(0.0039, 0.0082, 0.0150)
      & \cite{Wawer2005-us, Skarbinski2015-ni}\\
      $\tau_{U}$ & Transmissibility after acute phase
      & $\Triangular$(0.00077, 0.0014, 0.00251)
      & \cite{Hughes2012-so} \\
      $\varepsilon$ & Relative transmissibility with
      viral suppression & $\BetaPERT$(0.08, 0.002, 0.57)
      & \cite{Donnell2010-xo, attia_2009, wilson_2012, jia_2013,
        rodger_2016} \\
      $n$ & Coital acts per year & $\Uniform$(96, 108)
      & \cite{Wawer2005-us, Abdool_Karim2010-cm}\\
      $\alpha$ & Diagnosis rate & See eq.~(\ref{diagnosis_rate}) & --- \\
      $\phi$ & Treatment rate & See eq.~(\ref{treatment_rate}) & --- \\
      $\psi$ & Rate of relapse to untreated & See eq.~(\ref{relapse_rate})
      & --- \\
      $\zeta$ & Vaccination rate & See eq.~(\ref{vaccination_rate}) & --- \\
      \hline
      \vspace*{-2.5em}
    \end{tabularx}
    \caption{Model parameters. $\Triangular$, $\Uniform$, and
      $\BetaPERT$ are sampling distributions (\autoref{uncertainty}).
      For vaccine efficacy, $50\%$ was the baseline, and $30\%$ and
      $70\%$ were also used in a scenario analysis.}
    \label{model_param}
  \end{center}
\end{table}

People susceptible to HIV ($S$) progress to acute infection ($A$)
according to the force of infection ($\lambda$), which depends on the
HIV transmission rate, estimated from the country-specific incidence
and prevalence data (\autoref{model_fitting}), adjusted for the
current number of acutely infected, untreated or ineffectively
treated, and viral suppression PLHIV.  Acute infection is
characterized by high viral loads and lasts an average of 2.90 [95\%
confidence interval 1.23, 6.00]\;months\cite{Hollingsworth2008-iy}.
(We converted these durations to rates like
$1 / 2.90\;\text{month$^{-1}$} \approx 4.14\;\text{y$^{-1}$}$ and set
the parameter using random samples from
$\Triangular(2, 4.14, 9.6)\;\text{y$^{-1}$}$ to capture the
uncertainty in its value.  See \autoref{uncertainty} for the
definition of $\Triangular(a, b, c)$ and other distributions.)  We
assumed that people remain undiagnosed during their acute-infection
phase. After acute infection, people move into the chronic undiagnosed
class ($U$).  People in the undiagnosed class will be diagnosed at the
rate $\alpha$, which is determined by the country's current diagnosis
level and its diagnosis target (\autoref{targets}).  Diagnosed people
($D$) transition to the treated compartment ($T$) upon starting ART at
the rate $\phi$, dependent on the country's current treatment level
and its treatment target (\autoref{targets}).  After some time,
treated people then transition to the viral-suppression class ($V$),
at the rate $\gamma$.  We took $\gamma$ as a random sample from
$\Uniform(0.5, 1.5)\;\text{y$^{-1}$}$, that is, viral suppression
occurs after between 8 months and 2 years on
treatment\cite{Currie2009-yz}.  Disengagement from treatment moves
people from treated ($T$) and viral suppression ($V$) at the rate
$\psi$, dependent on the current level of viral suppression and the
country's target (\autoref{targets}), back to the diagnosed class
($D$).

PLHIV without viral suppression (compartments $U$, $D$, and $T$)
develop AIDS (compartment $W$) after an average duration of 10.4
years\cite{Morgan2002-cq}, i.e. with transition rate
$\sigma = 0.1064\;\text{y$^{-1}$}$.  People who have achieved viral
suppression ($V$) develop AIDS at rate $\theta$, slower than PLHIV
without viral suppression.  People with viral suppression are
estimated to live $5$--$8$\;y shorter lives than the non-HIV
infected\cite{Samji2013-kf, Unaids2014-ue}, which we quantified by
randomly sampling the parameter $\omega$ from
$\Uniform(5, 8)\;\text{y}$.  Assuming no relapse to untreated, the
duration of viral suppression is
\begin{equation}
  \frac{1}{\theta + \mu} = \frac{1}{\mu} - \omega - \frac{1}{\nu}.
\end{equation}
The term to the left of the equals sign is the time until leaving
viral suppression by either progression to AIDS (at rate $\theta$) or
background mortality (at rate $\mu$).  The term to the right of the
equals sign is the lifespan of the non-HIV infected, minus the
reduction in life due to having HIV with viral suppression, minus the
duration of the AIDS stage.  Solving for $\theta$ gives the rate of
progression to AIDS with viral suppression as
\begin{equation}
  \label{theta}
  \theta = \frac{1}{1/\mu - \omega - 1/\nu} - \mu.
\end{equation}

@@

The relative transmission rates of acute, unsuppressed, and suppressed
infected people are
\begin{equation}
  \label{betas}
  \begin{split}
    \beta_A &= 1 - (1 - \tau_A)^n,
    \\
    \beta_U &= 1 - (1 - \tau_U)^n,
    \\
    \beta_V &= 1 - (1 - \varepsilon \tau_U)^n,
  \end{split}
\end{equation}
where the $\tau$'s are the relative transmissibilities of people with
acute ($A$), unsuppressed ($U$ and $T$), and suppressed ($V$)
infections, $n$ is the annual number of coital acts, and $\varepsilon$
is the relative transmissibility with viral suppression.  For
transmissibility during the acute phase, we sampled $\tau_A$ from
$\Triangular(0.0039, 0.0082, 0.0150)$\cite{Wawer2005-us,
  Skarbinski2015-ni},
whereas, for transmissibility for unsuppressed people, we sampled
$\tau_U$ from
$\Triangular(0.00077, 0.0014, 0.00251)$\cite{Hughes2012-so}.  The
transmissibility of people with viral suppression relative to those
without suppression has been estimated at 0.08 (95\% confidence
interval 0.002, 0.57)\cite{Donnell2010-xo}: due to the extremely long
left tail, we chose to sample the relative transmissibility
$\varepsilon$ from $\BetaPERT(0.08, 0.002, 0.57)$ rather than a
$\Triangular$ distribution.  (See \autoref{uncertainty} for a full
description of $\BetaPERT(a, b, c)$.)  The annual number of coital
acts $n$ was sampled from
$\Uniform(96, 108)$\cite{Wawer2005-us, Abdool_Karim2010-cm}.  The
country-specific transmission rate $\eta$ was derived from UNAIDS
estimates of each country's longitudinal HIV prevalence and incidence
(\autoref{model_fitting}).

@@

The parametrized differential equations were solved numerically from
year 2015 to year 2035 using the LSODA routine\cite{odepack, scipy,
  medlock2016-git} and four effectiveness outcomes were computed from
the solutions.
\begin{description}
\item[New infections] New infections from 2015 to time $t$ are given
  by $Y(t)$.

@@

For 127 countries, we found sufficient published estimates and data to
parametrize our model.  Demographic data, including population growth
rate\cite{WorldBankpg}, death
rate\cite{World_Development_Indicators2013-ee}, and number of people
aged 15--49 years\cite{The_World_Bank2016-fd} were obtained from the
World Bank (Supplementary Table 2). Longitudinal HIV prevalence
(Supplementary Table 3) and incidence (Supplementary Table 4)
estimates for ages 15--49, spanning from as early as 1990 for some
countries to 2015, were primarily obtained from the AIDSinfo database
produced by UNAIDS\cite{Unaids2016-an}. Other sources included UNAIDS
Country Progress Reports\cite{Unaids2016-am} published from 2012 to
2016 and AIDS Data Hub\cite{AIDSdatahub-fg}, which compiles data from
UNAIDS, UNICEF, WHO, and the Asian Development Bank. These sources
were also used to inform the initial conditions for the model: the
number of people diagnosed with HIV, the number on ART, and the number
who have viral suppression or have been retained on treatment for at
least 12 months (Supplementary Table 5). Where estimates or data were
not available from these sources, alternative resources such as
peer-reviewed journal articles and country health ministry reports
were consulted.  See Supplementary Tables 2--5 for full information on
the sources used for each country.  These source tables are also
available in the public source-code repository\cite{medlock2016-git}.

@@

Using available historical country-level estimates of prevalence and
incidence (Supplementary Tables 3 and 4), and published estimates of
transmissibility\cite{Wawer2005-us, Donnell2010-xo, Hughes2012-so,
  Skarbinski2015-ni},
we derived an average country-specific transmission rate, which then
informs the acute, unsuppressed, and suppressed transmission rates.
This calibration method provided a recent estimate of country-specific
transmission, while using historical data to smooth out short-term
variations.

@@

To reflect the uncertainty in $\beta(t)$, we assumed that it follows a
lognormal distribution (\autoref{uncertainty}). For the parameters
$\mu$ and $\sigma^2$ of the lognormal distribution we used the
exponentially weighted mean and variance\cite{holt2004} over time
(with half-life 1\;y) of $\log \beta(t)$ for the year 2015.  That is,
\begin{equation}
  \label{lognormal_params}
  \begin{split}
    \mu &= \frac{\sum_{i = 1}^n  2^{t_i}
      \log \beta(t_i)}
    {\sum_{i = 1}^n 2^{t_i}},
    \\
    \sigma^2 &= \frac{\big(\sum_{i = 1}^n 2^{t_i}\big)
      \left[\sum_{i = 1}^n  2^{t_i}
        \big(\log \beta(t_i) - \mu\big)^2\right]}
    {\big(\sum_{i = 1}^n 2^{t_i}\big)^2
      - \sum_{i = 1}^n 2^{2 t_i}}.
  \end{split}
\end{equation}
A half-life of 1\;y means that the estimate of the transmission rate
from year $i$ is weighted half as much as the estimate from year
$i + 1$, a quarter as much as the estimate from year $i + 2$, and so
on.  This approach captured the recent country-level transmission
behavior while still allowing for the use of less-recent estimates to
smooth short-term variations (\autoref{transmission_rate}).

@@

Projections were generated for each country through model simulations
using 1000 random samples from published parameter distributions
(\autoref{model_param}) and the country-specific estimated
distribution of the transmission rate (\autoref{model_fitting}). The
parameters were sampled using Latin hypercube
sampling\cite{blower1994}.  We also used the variance reduction
technique of running different interventions with the same parameter
samples, i.e.  one set of sample parameter values is drawn, status quo
and intervention scenarios are all run with those sample values and
compared, and this process is repeated for each of the 1000 sample
sets of parameters\cite{shechter2006}.

The model outcomes were summarized using the median, 1st and 3rd
quartiles (i.e. 25th and 75th percentiles), and 5th and 95th
percentiles.  We calculated partial rank correlation coefficients
(PRCC)\cite{blower1994} to measure the independent effect of each
non-vaccine parameter on model projections of global new HIV
infections from 2015 to 2035 (Extended Data Fig.~\ref*{PRCCs}).

@@

\item[Beta-PERT] $\BetaPERT(a, b, c)$ is the Beta-PERT probability
  distribution\cite{malcom1959} with minimum $a$, mode $b$, and
  maximum $c$.  It has density function
  \begin{equation}
    \label{BetaPERT}
    f_{\BetaPERT}(x) =
    \begin{cases}
      \frac{(x - a)^{v - 1} (c - x)^{w - 1}}{(c - a)^{v + w - 2} B(v, w)}
      & \text{if $a \leq x \leq c$,}
      \\
      0 & \text{otherwise,}
    \end{cases}
  \end{equation}
  where
  \begin{equation}
    \begin{split}
      \mu &= \frac{a + \lambda b + c}{\lambda + 2},
      \\
      v &= \frac{(\mu - a)(2 b - a - c)}{(b - \mu) (c - a)},
      \\
      w &= \frac{v (a - \mu)}{\mu - c},
    \end{split}
  \end{equation}
  and $B(v, w)$ is the standard beta function\cite{davis1972}.
  % \cite[\S6.2]{davis1972}
  We used the standard value of $\lambda = 4$.

@@

% \printbibliography
\bibliography{supplementary_information}
\bibliographystyle{naturemag}
